Seasonal Variations in Marriage License Applications

### 2024-09-27 paper/paper.qmd
comma removal
final checks

---
title: "Seasonal Variations in Marriage License Applications"
subtitle: "An Exploratory Analysis with Open Data Toronto"
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at: https://github.com/fatimahsy/Pandemic-Marriages-.git."
date: today
date-format: long
abstract: "Marriage License data from Toronto between 2011 and 2024 was analysed to explore trends in marriage licience number and the influence of emotionally significant dates like valentines day & christmas on these numbers. Couples are more practical in choosing wedding dates, with significantl higher numbers of marriage liciences issued during the summer months.These results hilighthow societal traditions around marriage are shaped more by practical considerations than romantic ideals. Understanding these patterns can help city officials anticipate changes in demand for marriage-related services and allocate resources accordingly, ensuring that civil centers are prepaered for these seasonal shifts."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

# Introduction
Marriage decisions are often thought to be driven by deeply emotional and symbolic impulses, such as getting married on Valentine's Day or during the Christmas holidays. However, this paper finds that practical considerations like favorable weather and scheduling play a more significant role in choosing wedding dates. By examining marriage license data in Toronto from 2011 to 2023, this paper looks at trends in key periods such as February, the summer months, and December showing how couples tend to prioritize convenience over sentimentality, particularly favoring summer months for weddings.

@@

We use data analysis and different visualizations to examine marriage license trends from 2011 to 2023, with a focus on key seasons such as summer, Christmas, and Valentine’s Day. While marriage license issuance declined significantly during the pandemic, these symbolic dates continued to show small spikes, indicating that emotional or symbolic significance still plays a role in some marriage decisions. However, the data leans more towards suggesting that practical considerations, such as weather and scheduling, are the primary drivers of marriage planning, with the summer months consistently seeing the highest activity.

Marriage trends are a very good indicator of changes in society, reflecting broader shifts in demographics, economic circumstances, and changing cultural preferences. Understanding how external factors like seasonal holidays and different disruptions to the overall society influence marriage decisions can offer us valuable insights into human behavior. By identifying consistent seasonal patterns, this analysis can help wedding-related businesses (venues, caters, DJ's, etc.)anticipate periods of higher demand and better allocate resource. Additionally, civil centers can use this data to optimise staffing in order to increase efficiency.

@@

For further cleaning, the `janitor` (@citejanitor) package was used and the `knitr` was used too.

@@

Using Data about marriage licenses in Toronto, this paper tries to to see if there is a visible pattern that can suggest that people get married based on emotions or practicality. It focuses on emotionally significant dates like Valentines day, Christmas, and summer. The findings reveal that practical considerations, like weather and guest availability tend to outweigh emotional or sentimental motivations as seen in  @sec-results. There are consistent increases during the summer months compared to the relatively lower numbers on Valentine's Day or during the Winter.

@@

One of the clearest patters shown from the data is the overwhelming preference for summer months when it comes t getting married. Across all years, these moths far exceed other months. This is largely attributed to the fact that during summer, logistics are much easier to work out. People are able to have weddings outside which is usually preferred. People also do not have to worry about bad weather like snow storms or heavy wind. Having a wedding during the summer also makes pictures look better since the sun is shining pretty well. In terms of guest availability, summer is usually optimal because there is no school, and many people take time off work.

Even thought dates like valentines day hold cultural and romantic significance, they do not show a corresponding increase in marriage activity. This indicates that while couples may enjoy celebrating love or symbolic days, their wedding planning is more practically driven by factors like convenience, weather, and timing.

While the findings offer important insights into marriage license trends, there are a few limitations. First, the analysis is restricted to marriage license data which does not account for actual ceremonies or couples' personal motivations for choosing specific dates. While the data clearly shows summer dates are more popular, it cannot directly explain why couples make these decisions or weather other factors like venue availability or budget considerations play a role. Additionally, the dataset does not include demographic information like age, income, race or cultural background which could help provide a better understanding of why certain groups might favour particular wedding dates.

@@

\`\`\`

>>>>

---
title: "Seasonal Variations in Marriage License Applications"
subtitle: "An Exploratory Analysis with Open Data Toronto"
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at: https://github.com/fatimahsy/Pandemic-Marriages-.git."
date: today
date-format: long
abstract: "Marriage License data from Toronto between 2011 and 2024 was analysed to explore trends in marriage licience numbers and the influence of emotionally significant dates like valentines day & christmas on these numbers. Couples are more practical in choosing wedding dates, with significantly higher numbers of marriage liciences issued during the summer months.These results hilight how societal traditions around marriage are shaped more by practical considerations than romantic ideals. Understanding these patterns can help city officials anticipate changes in demand for marriage-related services and allocate resources accordingly, ensuring that civil centers are prepaered for these seasonal shifts."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

# Introduction
Marriage decisions are often thought to be driven by deeply emotional and symbolic impulses, such as getting married on Valentine's Day or during the Christmas holidays. However, this paper finds that practical considerations like favorable whether and scheduling play a more significant role in choosing wedding dates. By examining marriage license data in Toronto from 2011 to 2023, this paper looks at trends in key periods such as February, the summer months, and December showing how couples tend to prioritize convenience over sentimentality, particularly favoring summer months for weddings.

@@

We use data analysis and different visualizations to examine marriage license trends from 2011 to 2023, with a focus on key seasons such as summer, Christmas, and Valentine’s Day. While marriage license issuance declined significantly during the pandemic, these symbolic dates continued to show small spikes, indicating that emotional or symbolic significance still plays a role in some marriage decisions. However, the data leans more towards suggesting that practical considerations, such as whether and scheduling, are the primary drivers of marriage planning, with the summer months consistently seeing the highest activity.

Marriage trends are a very good indicator of changes in society, reflecting broader shifts in demographics, economic circumstances, and changing cultural preferences. Understanding how external factors like seasonal holidays and different disruptions to the overall society influence marriage decisions can offer us valuable insights into human behavior. By identifying consistent seasonal patterns, this analysis can help wedding-related businesses (venues, caterers, DJ's, etc.)anticipate periods of higher demand and better allocate resource. Additionally, civil centers can use this data to optimise staffing in order to increase efficiency.

@@

For further cleaning, the `janitor` (@citejanitor) package was used and the `knitr` package was used too.

@@

Using Data about marriage licenses in Toronto, this paper tries to see if there is a visible pattern that can suggest that people get married based on emotions or practicality. It focuses on emotionally significant dates like Valentines day, Christmas, and summer. The findings reveal that practical considerations, like whether and guest availability tend to outweigh emotional or sentimental motivations as seen in  @sec-results. There are consistent increases during the summer months compared to the relatively lower numbers on Valentine's Day or during the Winter.

@@

One of the clearest patterns shown from the data is the overwhelming preference for summer months when it comes to getting married. Across all years, these months far exceed other months. This is largely attributed to the fact that during summer, logistics are much easier to work out. People are able to have weddings outside which is usually preferred. People also do not have to worry about bad whether like snow storms or heavy wind. Having a wedding during the summer also makes pictures look better since the sun is shining pretty well. In terms of guest availability, summer is usually optimal because there is no school, and many people take time off work.

Even though dates like valentines day hold cultural and romantic significance, they do not show a corresponding increase in marriage activity. This indicates that while couples may enjoy celebrating love or symbolic days, their wedding planning is more practically driven by factors like convenience, whether, and timing.

While the findings offer important insights into marriage license trends, there are a few limitations. First, the analysis is restricted to marriage license data which does not account for actual ceremonies or couples' personal motivations for choosing specific dates. While the data clearly shows summer dates are more popular, it cannot directly explain why couples make these decisions or whether other factors like venue availability or budget considerations play a role. Additionally, the dataset does not include demographic information like age, income, race or cultural background which could help provide a better understanding of why certain groups might favour particular wedding dates.

@@
